Chemical potential of liquids and mixtures via Adaptive Resolution Simulation

### 2014-05-10 writeDown/pub.jcp/chem-pot-jcp.tex
add the prove for the mixture

% LaTex
%\documentclass[aps,twocolumn]{revtex4-1}
\documentclass[a4paper,preprint,unsortedaddress]{revtex4-1}
%\documentclass[a4paper]{article}
\usepackage{amsmath,amssymb,amsfonts,latexsym,graphicx}
\usepackage[dvips]{color}
\usepackage{indentfirst}

@@

\def\pd#1#2{\frac{\partial #1}{\partial #2}}
\renewcommand{\vec}[1]{\mbox{\boldmath \small $#1$}}
\newcommand{\concen}{C}
\setlength\parindent{2em} 
\newcommand{\recheck}[1]{{\color{red} #1}}
\newcommand{\redc}[1]{{\color{red} #1}}            
\newcommand{\bluec}[1]{{\color{blue} #1}}            
\newcommand{\vect}[1]{\textbf{\textit{#1}}}
\newcommand{\dof}{{\textrm{DOF}}}
\newcommand{\AT}{{\textrm{{AT}}}}
\newcommand{\EX}{{\textrm{EX}}}
\newcommand{\CG}{{\textrm{CG}}}
\newcommand{\HY}{{\Delta}}
\newcommand{\rdf}{{\textrm{rdf}}}
\newcommand{\thf}{{\textrm{th}}}
\newcommand{\res}{{\textrm{rep}}}
\newcommand{\ext}{{\textrm{extra}}}
\newcommand{\exc}{{\textrm{ex}}}
\newcommand{\thermo}{{\textrm{Q}}}
\newcommand{\hadress}{{\textrm{H}}}
\newcommand{\dadress}{{\textrm{D}}}

@@

\bluec{
  Unified of terminology:
  \begin{itemize}
  \item interface region $\rightarrow$ transition region
  \item hybrid region $\rightarrow$ transition region
  \item define COM, change all ``center of mass'' to COM
  \item change the title of sec.III
  \item do not be too mean, we cite BOTH of their PRL H-Adress paper...
  \item some times, it is called GC-AdResS, sometimes original AdResS, sometimes standard AdResS, sometimes force-interpolation AdResS....
  \item rewrite eq. 4, see if it is OK for you
  \end{itemize}
}

@@

In Ref.~\cite{prx} it has been shown that the chemical potential of the atomistic and coarse-grained resolution are related by the following formula: 
\begin{equation}
\mu_{\CG}=\mu_{\AT}+\omega_{\thf}+\omega_{\thermo}
\label{mu}
\end{equation}
with $\mu_{\CG}$ the chemical potential of the coarse-grained system (in GC-AdResS this corresponds to a liquid of generic spheres), $\mu_{\AT}$  the chemical potential of the atomistic system, $\omega_{\thf}=\int_{\Delta}{\vect F}_{\thf}({\vect r})d{\vect r}$ the work of the thermodynamic force in the transition region, $\omega_{\thermo}$ the work/heat provided by the thermostat in order to slowly equilibrate the inserted/removed degrees of freedom in the transition region. $\omega_{\thermo}$ is composed by two parts, one, called $\omega_{\ext}$, which compensates the dissipation of energy due to the change of interactions in $\Delta$, and another, $\omega_{\dof}$, which is related to the equilibration of the reinserted/removed degrees of freedom (rotational and vibrational). {While the determination of $\omega_{\dof}$ is not required for the our final aim (that is the calcualtion of the excess chemical potential, as explain later on), the calculation of $\omega_{\thermo}$ is very relevant. However this calculation is not straightforward and we have proposed to introduce an auxiliary Hamiltonian approach where the coarse-grained and atomistic potential are interpolated, and not the forces as in AdResS. Next, we impose that the Hamiltonian system must have the same thermodynamic equilibrium of the original force-based AdResS system; this is done by introducing a thermodynamic force in the auxiliary Hamiltonian approach, which, at the target temperature, keeps the density of particles across the system as in AdResS. \recheck{In the auxiliary Hamiltonian approach we have the same equilibrium as the original adaptive (and full atomistic) system  and the difference between the work of the original thermodynamic force and the work of the thermodynamic force calculated in the Hamiltonian approach gives $\omega_{\ext}$ (further details about this point are given in the Appendix)}. Moreover we have proven numerically, for the case of liquid water, that
  $\omega_{\ext}=\int_{\Delta}\nabla w(\vect r)\langle w( U^{\AT}-U^{\CG})\rangle_{\vect r} d{\vect r}$,
  where $U_{AT}$ and $U_{CG}$ are the atomistic and coarse-grained potential. 
  \recheck{In the next section of this work we show analytically that the formula above is exact (at least) at the first order w.r.t. the probability distribution of the system as defined in Eq.\ref{order}}.

@@

\section{Analytic Deriviation of the $w_{\ext}$}
{In this section we derive analytically the equivalence:
  $\omega_{\ext}=\int_{\Delta}\nabla_{\vect r} w(\vect r) \langle w(U^{\AT}-U^{\CG})\rangle_{\vect r} d{\vect r}$ and define its limitations.
We define a potential coupling the atomistic and coarse-grained resolution as the spatial interpolation of the atomistic and coarse-grained potential, as done instead for the forces in the standard AdResS:
\begin{align}
  U =
  \sum_{i<j} w(\vect r_i) w(\vect r_j)
  U^\AT_{i,j}
  +
  \sum_{i<j}[1- w(\vect r_i) w(\vect r_j)]
  U^\CG_{i,j},
\end{align}
where $U^\AT_{i,j}$ and $U^\CG_{i,j}$ are the atomistic and coarse-grained interaction potential between molecule $i$ and $j$, respectively, defined by
\begin{align}\label{eqn:pot-interpol}
  U^\AT_{i,j} =
  \sum_{\alpha\in i}\sum_{\beta\in j} U^\AT(\vect r_\alpha - \vect r_\beta),\quad
  U^\CG_{i,j} =
  U^\CG(\vect r_i - \vect r_j),
\end{align}
where index $\alpha$ and $\beta$ denotes the atom position of the corresponding molecule.
The COM of the molecule is defined as:
\begin{align}
  \vect r_i = \sum_{\alpha\in i} \frac{m_\alpha}{\sum_{\alpha\in i} m_\alpha} \vect r_\alpha,
\end{align}
where $m_\alpha$ is the mass of atom $\alpha$.
\recheck{The potential interpolation~\eqref{eqn:pot-interpol} provides an auxiliary Hamiltonian to the AdResS system, and
the
corresponding intermolecular force is given by:}
\begin{align}\label{eqn:interpol-f}
  \vect F_{i,j} =
  w(\vect r_i) w(\vect r_j)
  \vect F^\AT_{i,j}
  +
  [1- w(\vect r_i) w(\vect r_j)]
  \vect F_{i,j}^\CG
  -
  \nabla_{\vect r} w(\vect r_i) w(\vect r_j)
  (U^\AT_{i,j} - U^\CG_{i,j}).
\end{align}
\recheck{We refer to the AdResS simulation using force scheme~\eqref{eqn:interpol-f} as
  auxiliary Hamiltonian AdResS, and all properties of this approach will be added a superscript ``H''}.
We define the force of changing representation by
\begin{align}\label{eqn:frep}
  \vect F_{\res,i} = 
  \sum_j \nabla_{\vect r} w(\vect r_i) w(\vect r_j)
  (U^\AT_{i,j} - U^\CG_{i,j}).
\end{align}
%An important fact is that the force of changing representation is not
%symmetric w.r.t molecule $\alpha$ and $\beta$, therefore, the Newton's
%action-reaction law (momentum conservation) does not hold anymore for
%the potential interpolation approach. On the other hand, the force
%interpolation scheme does not preserve
%energy, while the potential interpolation
%approach has the advantage of energy conservation. In sum, both the
%force and potential interpolation approaches do not hold all physical
%conservation laws, because the hybrid region is artificially
%designed. Therefore, we will take the advantage of both and
%develop the thermodynamic relation (between the atomistic and
%coarse-grained resolutions)  for them at the same time.\\

@@

It is straightforward to show that \recheck{HAN, here it must be shown} \bluec{No, who cannot derive it should be sent back to the graduate course of thermodynamics.} 

@@

  = \int_\HY d \vect r\,\nabla_{\vect r}\omega(\vect r)\,
  \big\langle \omega(\vect r')\, [U^\AT(\vect r-\vect r') - U^\CG(\vect r-\vect r')]\big\rangle_{\vect r'; \vect r}

@@

\recheck{The average is taken over all possible positions of the second molecule, provided the position of the first molecule
in the pairwise interaction}. In case of molecules that contain

@@

conformations of the molecules. 

@@

\bluec{here since we have the auxiliary Hamiltonian, then the system can be treated by the conclusions
  that can be found in standard thermodynamic text books. If the referee does not understand, he should
be sent back to the graduate cource. }

@@

From Eq.~\eqref{eqn:hd-rel} and \eqref{eqn:hd-rel-2}, we
find the extra work of the thermostat in GC-AdResS being identical to the work of changing
representation of the Hamiltonian auxiliary approach:
\begin{align}\label{eqn:wextra}
  \omega_\ext = \omega_\res,
\end{align}
which basically proves the statement at the beginning of this section.  The ensemble
average on the R.H.S.~of Eq.~\eqref{eqn:w-rep} is performed in the
ensemble of the system treated with the potential interpolation approach, and the question is if the ensemble average is equivalent if it is performed in the simulation where the force
interpolation approach is used.  It is obvious that the phase spatial probability distribution correponding to the system treated with the potential interpolation is consistent with the force
interpolation at least up to the 1st order as defined in Eq.\ref{order}.  It is also
possbile to systematically obtain equivalence in the ensemble average operation  at higher orders of accuracy of the probability distribution, as, for example, it is done for the radial distribution function in Ref.~\cite{jctchan}.
However, here we do not consider higher order corrections, because it
has been numerically shown that actually the ensemble average of
$\vect F_\res$ dose not depend on in which ensemble it is
calculated~\cite{prx}. Therefore, we use Eq.~\eqref{eqn:wextra} to
calculate $\omega_\ext$, and measure the ensemble average by the
standard AdResS.  Since in the computer implementation, the CG
molecules also keep the atomistic degrees of freedom eventhough they
are in the CG region, therefore, the kinetic part of $\mu_\AT$ and
$\mu_\CG$ are identical, and $\omega_\dof$ vanishes. Therefore,
by inserting Eq.~\eqref{eqn:wextra} into \eqref{eqn:mueq-d}, we have
\begin{align}
  \mu_\AT^\exc = \mu_\CG^\exc - \int_\HY d\vect r \vect F_\thf(\vect r)
-\int_\HY d \vect r\,\nabla_{\vect r}\omega(\vect r)\,
\big\langle \omega(\vect r')\, [U^\AT(\vect r-\vect r') - U^\CG(\vect r-\vect r')]\big\rangle_{\vect r'; \vect r}
% -\int_\HY dr \nabla\omega(r)\,\big\langle \omega(r')\, [U^\AT(r-r') - U^\CG(r-r')]\big\rangle_{r'}
\end{align}
In the next section we apply the procedure to the calculation of the excess chemical potential of several liquids and mixtures.

\section{Results and Discussion}
We have calculated $\mu^{ex}$ for different liquids and mixtures, choosing cases which are representative of a large class of systems. Hydrophobic solvation in methane/water and in ethane/water mixtures, hydrophilic solvation in urea/water, a balance of both in water/tert-Butyl alcohol (TBA) mixture, other liquids, e.g. pure methanol and DMSO (and their mixtures with water), non aqueous mixtures in TBA/DMSO and alkane liquids such as methane, ethane and propane. Moreover, systems as water/urea are commonly used as cosolvent of biological molecules \cite{nico-debashish} while systems as tert-Butyl alcohol/water play a key role in modern technology \cite{irata}, thus they are of high interest {\it per se}. All technical details of each simulation are presented in the Appendix.
Results are reported in Table \ref{table} where the comparison with values obtained using full atomistic TI and available experiments, at the same concentrations, of our calculation is made; in our previous work we have already shown that value of the chemical potential of liquid water obtained with IPM is well reproduced by GC-AdResS, however the computational cost of IMP was very large, thus we do not consider calculations done with IPM in this paper. 
The agreement with full atomistic TI simulations is satisfactory in all cases, and thus it proves the solidity of GC-AdResS in describing the essential thermodynamics of a large class of systems.
We also compare the obtained values with those available in literature \cite{vang,nico}. Although the concentration of the minor component in the mixtures that we consider, is higher than the concentrations considered in Refs.\cite{vang,nico}, we are anyway in the very dilute regime and thus the chemical potential should not change in a significant way; we have verified  such a supposed consistency. 
The chemical potential of $k$-th liquid's component in a mixture is calculated as:
\begin{equation}
  \mu^{\exc,k}_{\AT}=
  \mu^{\exc,k}_{\CG}-
  \int_{\Delta}{\vect F}^k_{\thf}({\vect r})d{\vect r}-
  \int_{\Delta}\nabla_{\vect r} w(\vect r) \langle w(U^{\AT}-U^{\CG})\rangle_{\vect r, k} d{\vect r}
% \mu^{ex,i}_{AT}=\mu^{ex,i}_{CG}-\int_{\Delta}{F}^{i}_{th}({\vect r})d{\vect r}-\int_{\Delta}\langle w\nabla_{x_{i}} w(x)(V_{AT}-V_{CG})\rangle_{\vect r}d{\vect r}
\label{icomp}
\end{equation}
where ${F}^{k}_{\thf}(x)$ is the thermodynamic force applied to the molecules of the $k$-th component; this assures that, at the given concentration, the density of molecules of species $k$, in the transition region, is equivalent to the density of the same liquid's component in a reference full atomistic simulation.
\recheck{The ensemble evarage is taken over the position of the second molecule, provided that the first moelcule is of species $k$, and located at postion $\vect r$}.
% $\nabla_{x_{i}} w(x)$ is the gradient along $x$ (in a rectangular box) applied to the switching function $w(x)$ in the COM of molecules of component $i$.

@@

>>>>

% LaTex
%\documentclass[aps,twocolumn]{revtex4-1}
\documentclass[a4paper,preprint,unsortedaddress]{revtex4-1}
%\documentclass[a4paper]{article}
\usepackage{amsmath,amssymb,amsfonts,latexsym}
\usepackage[dvips]{graphicx}
\usepackage{color}
\usepackage{indentfirst}

@@

\def\pd#1#2{\frac{\partial #1}{\partial #2}}
\renewcommand{\vec}[1]{\mbox{\boldmath \small $#1$}}
\newcommand{\concen}{C}
\setlength\parindent{2em} 
\newcommand{\recheck}[1]{{\color{red} #1}}
\newcommand{\redc}[1]{{\color{red} #1}}            
\newcommand{\bluec}[1]{{\color{blue} #1}}            
\newcommand{\vect}[1]{\textbf{\textit{#1}}}
\newcommand{\dof}{{\textrm{DOF}}}
\newcommand{\AT}{{\textrm{{AT}}}}
\newcommand{\EX}{{\textrm{EX}}}
\newcommand{\CG}{{\textrm{CG}}}
\newcommand{\HY}{{\Delta}}
\newcommand{\rdf}{{\textrm{rdf}}}
\newcommand{\thf}{{\textrm{th}}}
\newcommand{\res}{{\textrm{rep}}}
\newcommand{\ext}{{\textrm{extra}}}
\newcommand{\exc}{{\textrm{ex}}}
\newcommand{\thermo}{{\textrm{Q}}}
\newcommand{\hadress}{{\textrm{H}}}
\newcommand{\dadress}{{\textrm{D}}}
\newcommand{\typea}{{\textrm{A}}}
\newcommand{\typeb}{{\textrm{B}}}

@@

@@

In Ref.~\cite{prx} it has been shown that the chemical potential of the atomistic and coarse-grained resolution are related by the following formula: 
\begin{equation}
\mu_{\CG}=\mu_{\AT}+\omega_{\thf}+\omega_{\thermo}
\label{mu}
\end{equation}
with $\mu_{\CG}$ the chemical potential of the coarse-grained system (in GC-AdResS this corresponds to a liquid of generic spheres), $\mu_{\AT}$  the chemical potential of the atomistic system, $\omega_{\thf}=\int_{\Delta}{\vect F}_{\thf}({\vect r})d{\vect r}$ the work of the thermodynamic force in the transition region, $\omega_{\thermo}$ the work/heat provided by the thermostat in order to slowly equilibrate the inserted/removed degrees of freedom in the transition region. $\omega_{\thermo}$ is composed by two parts, one, called $\omega_{\ext}$, which compensates the dissipation of energy due to the \bluec{non-conservative interactions} in $\Delta$, and another, $\omega_{\dof}$, which is related to the equilibration of the reinserted/removed degrees of freedom (rotational and vibrational). {While the determination of $\omega_{\dof}$ is not required for the our final aim (that is the calcualtion of the excess chemical potential, as explain later on), the calculation of $\omega_{\thermo}$ is very relevant. However this calculation is not straightforward and we have proposed to introduce an auxiliary Hamiltonian approach where the coarse-grained and atomistic potential are interpolated, and not the forces as in AdResS. Next, we impose that the Hamiltonian system must have the same thermodynamic equilibrium of the original force-based AdResS system; this is done by introducing a thermodynamic force in the auxiliary Hamiltonian approach, which, at the target temperature, keeps the density of particles across the system as in AdResS. \recheck{In the auxiliary Hamiltonian approach we have the same equilibrium as the original adaptive (and full atomistic) system  and the difference between the work of the original thermodynamic force and the work of the thermodynamic force calculated in the Hamiltonian approach gives $\omega_{\ext}$ (further details about this point are given in the Appendix)}. Moreover we have proven numerically, for the case of liquid water, that
  $\omega_{\ext}=\int_{\Delta}\nabla w(\vect r)\langle w( U^{\AT}-U^{\CG})\rangle_{\vect r} d{\vect r}$,
  where $U_{AT}$ and $U_{CG}$ are the atomistic and coarse-grained potential. 
  \recheck{In the next section of this work we show analytically that the formula above is exact (at least) at the first order w.r.t. the probability distribution of the system as defined in Eq.\ref{order}}.

@@

\section{Analytic Deriviation of the $w_{\ext}$}\label{sec:tmp4}
{In this section we derive analytically the equivalence:
  $\omega_{\ext}=\int_{\Delta}\nabla_{\vect r} w(\vect r) \langle w(U^{\AT}-U^{\CG})\rangle_{\vect r} d{\vect r}$ and define its limitations.
We define a potential coupling the atomistic and coarse-grained resolution as the spatial interpolation of the atomistic and coarse-grained potential, as done instead for the forces in the standard AdResS:
\begin{align}
  U =
  \sum_{i<j} w(\vect r_i) w(\vect r_j)
  U^\AT_{i,j}
  +
  \sum_{i<j}[1- w(\vect r_i) w(\vect r_j)]
  U^\CG_{i,j},
\end{align}
where $U^\AT_{i,j}$ and $U^\CG_{i,j}$ are the atomistic and coarse-grained interaction potential between molecule $i$ and $j$, respectively, defined by
\begin{align}\label{eqn:pot-interpol}
  U^\AT_{i,j} =
  \sum_{\alpha\in i}\sum_{\beta\in j} U^\AT(\vect r_\alpha - \vect r_\beta),\quad
  U^\CG_{i,j} =
  U^\CG(\vect r_i - \vect r_j),
\end{align}
where index $\alpha$ and $\beta$ denotes the atom position of the corresponding molecule.
The COM of the molecule is defined as:
\begin{align}
  \vect r_i = \sum_{\alpha\in i} \frac{m_\alpha}{\sum_{\alpha\in i} m_\alpha} \vect r_\alpha,
\end{align}
where $m_\alpha$ is the mass of atom $\alpha$.
\recheck{The potential interpolation~\eqref{eqn:pot-interpol} provides an auxiliary Hamiltonian to the AdResS system, and
the
corresponding intermolecular force is given by:}
\begin{align}\label{eqn:interpol-f}
  \vect F_{i,j} =
  w(\vect r_i) w(\vect r_j)
  \vect F^\AT_{i,j}
  +
  [1- w(\vect r_i) w(\vect r_j)]
  \vect F_{i,j}^\CG
  -
  \nabla_{\vect r} w(\vect r_i) w(\vect r_j)
  (U^\AT_{i,j} - U^\CG_{i,j}).
\end{align}
\recheck{We refer to the AdResS simulation using force scheme~\eqref{eqn:interpol-f} as
  auxiliary Hamiltonian AdResS, and all properties of this approach will be added a superscript ``H''}.
We define the force of changing representation by
\begin{align}\label{eqn:frep}
  \vect F_{\res,i} = 
  \sum_j \nabla_{\vect r} w(\vect r_i) w(\vect r_j)
  (U^\AT_{i,j} - U^\CG_{i,j}).
\end{align}
%An important fact is that the force of changing representation is not
%symmetric w.r.t molecule $\alpha$ and $\beta$, therefore, the Newton's
%action-reaction law (momentum conservation) does not hold anymore for
%the potential interpolation approach. On the other hand, the force
%interpolation scheme does not preserve
%energy, while the potential interpolation
%approach has the advantage of energy conservation. In sum, both the
%force and potential interpolation approaches do not hold all physical
%conservation laws, because the hybrid region is artificially
%designed. Therefore, we will take the advantage of both and
%develop the thermodynamic relation (between the atomistic and
%coarse-grained resolutions)  for them at the same time.\\

@@

It is straightforward to show that
% \recheck{HAN, here it must be shown} \bluec{No, who cannot derive it should be sent back to the graduate course of thermodynamics.} 

@@

  = \int_\HY d \vect r\,\nabla_{\vect r}w(\vect r)\,
  \big\langle w(\vect r')\,
  [\sum_{\alpha,\beta}U^\AT(\vect r_\alpha-\vect r'_\beta) - U^\CG(\vect r-\vect r')]\big\rangle_{\vect r'; \vect r}

@@

\bluec{The average is taken over all
  possible positions of the second molecule (i.e.~$\vect r'$),
  provided the position of the first molecule (i.e.~$\vect r$)
  in the pairwise interaction.
In case of molecules that contain

@@

conformations in the atomistic resolution.}

@@

% \bluec{here since we have the auxiliary Hamiltonian, then the system can be treated by the conclusions
%   that can be found in standard thermodynamic text books. If the referee does not understand, he should
% be sent back to the graduate cource. }

@@

From Eq.~\eqref{eqn:hd-rel} and \eqref{eqn:hd-rel-2}, we
find the extra work of the thermostat in GC-AdResS being identical to the work of changing
representation of the Hamiltonian auxiliary approach:
\begin{align}\label{eqn:wextra}
  \omega_\ext = \omega_\res,
\end{align}
which basically proves the statement at the beginning of this section.  The ensemble
average on the R.H.S.~of Eq.~\eqref{eqn:w-rep} is performed in the
ensemble of the system treated with the potential interpolation approach, and the question is if the ensemble average is equivalent if it is performed in the simulation where the force
interpolation approach is used.  It is obvious that the phase spatial probability distribution correponding to the system treated with the potential interpolation is consistent with the force
interpolation at least up to the 1st order as defined in Eq.\ref{order}.  It is also
possbile to systematically obtain equivalence in the ensemble average operation  at higher orders of accuracy of the probability distribution, as, for example, it is done for the radial distribution function in Ref.~\cite{jctchan}.
However, here we do not consider higher order corrections, because it
has been numerically shown that actually the ensemble average of
$\vect F_\res$ dose not depend on in which ensemble it is
calculated~\cite{prx}. Therefore, we use Eq.~\eqref{eqn:wextra} to
calculate $\omega_\ext$, and measure the ensemble average by the
standard AdResS.  Since in the computer implementation, the CG
molecules also keep the atomistic degrees of freedom eventhough they
are in the CG region, therefore, the kinetic part of $\mu_\AT$ and
$\mu_\CG$ are identical, and $\omega_\dof$ vanishes. Therefore,
by inserting Eq.~\eqref{eqn:wextra} into \eqref{eqn:mueq-d}, we have
\begin{align}\label{eqn:tmp24}
  \mu_\AT^\exc = \mu_\CG^\exc - \int_\HY d\vect r \vect F_\thf(\vect r)
-\int_\HY d \vect r\,\nabla_{\vect r}w(\vect r)\,
\big\langle w(\vect r')\, [\sum_{\alpha,\beta}U^\AT(\vect r_\alpha-\vect r'_\beta) - U^\CG(\vect r-\vect r')]\big\rangle_{\vect r'; \vect r}
% -\int_\HY dr \nabla\omega(r)\,\big\langle \omega(r')\, [U^\AT(r-r') - U^\CG(r-r')]\big\rangle_{r'}
\end{align}
In the next section we apply the procedure to the calculation of the excess chemical potential of several liquids and mixtures.

\section{Results and Discussion}
We have calculated $\mu^{ex}$ for different liquids and mixtures, choosing cases which are representative of a large class of systems. Hydrophobic solvation in methane/water and in ethane/water mixtures, hydrophilic solvation in urea/water, a balance of both in water/tert-Butyl alcohol (TBA) mixture, other liquids, e.g. pure methanol and DMSO (and their mixtures with water), non aqueous mixtures in TBA/DMSO and alkane liquids such as methane, ethane and propane. Moreover, systems as water/urea are commonly used as cosolvent of biological molecules \cite{nico-debashish} while systems as tert-Butyl alcohol/water play a key role in modern technology \cite{irata}, thus they are of high interest {\it per se}. All technical details of each simulation are presented in the Appendix.
Results are reported in Table \ref{table} where the comparison with values obtained using full atomistic TI and available experiments, at the same concentrations, of our calculation is made; in our previous work we have already shown that value of the chemical potential of liquid water obtained with IPM is well reproduced by GC-AdResS, however the computational cost of IMP was very large, thus we do not consider calculations done with IPM in this paper. 
The agreement with full atomistic TI simulations is satisfactory in all cases, and thus it proves the solidity of GC-AdResS in describing the essential thermodynamics of a large class of systems.
We also compare the obtained values with those available in literature \cite{vang,nico}. Although the concentration of the minor component in the mixtures that we consider, is higher than the concentrations considered in Refs.\cite{vang,nico}, we are anyway in the very dilute regime and thus the chemical potential should not change in a significant way; we have verified  such a supposed consistency. 
The chemical potential of $k$-th liquid's component in a mixture is calculated as:
\begin{equation}
  \mu^{\exc,k}_{\AT}=
  \mu^{\exc,k}_{\CG}-
  \int_{\Delta}{\vect F}^k_{\thf}({\vect r})d{\vect r}-
  \int_{\Delta}\nabla_{\vect r} w(\vect r) \langle w(U^{\AT}-U^{\CG})\rangle_{\vect r, k} d{\vect r}
% \mu^{ex,i}_{AT}=\mu^{ex,i}_{CG}-\int_{\Delta}{F}^{i}_{th}({\vect r})d{\vect r}-\int_{\Delta}\langle w\nabla_{x_{i}} w(x)(V_{AT}-V_{CG})\rangle_{\vect r}d{\vect r}
\label{eqn:tmp25}
\end{equation}
where ${F}^{k}_{\thf}(x)$ is the thermodynamic force applied to the molecules of the $k$-th component; this assures that, at the given concentration, the density of molecules of species $k$, in the transition region, is equivalent to the density of the same liquid's component in a reference full atomistic simulation.
\recheck{The ensemble evarage is taken over the position of the second molecule, provided that the first moelcule is of species $k$, and located at postion $\vect r$}.
% $\nabla_{x_{i}} w(x)$ is the gradient along $x$ (in a rectangular box) applied to the switching function $w(x)$ in the COM of molecules of component $i$.

@@

\section{Extending the chemical potential expression to multi-component system}

In this section we explain how to extend the chemical potential
expresstion Eq.~\eqref{eqn:tmp24} for pure system to multi-component
system, i.e.~\eqref{eqn:tmp25}.
For simplicity and without lost of generality, we assume that the system
is composed by two components, and the number of molecules are
$N^\typea$ and $N^\typeb$, respectively. The superscript ``$\typea$'' and ``$\typeb$'' throughout this section 
denotes the index of the components.  We further denote the number of 
molecule $\typea$ in the atomistic, transition, and coarse-grained regions by
$N^\typea_1$, $N^\typea_2$ and $N^\typea_3$, respectively. While those for type $\typeb$ molecules
are $N^\typeb_1$, $N^\typeb_2$ and $N^\typeb_3$, respectively.  By
assuming the transition region is infinitely small comparing with the
atomistic and coarse-grained regions, we have the following
constrains:
\begin{align}
  V &= V_1 + V_3\\
  N &= N^\typea + N^\typeb\\
  N^\typea &= N^\typea_1 + N^\typea_3\\
  N^\typeb &= N^\typeb_1 + N^\typeb_3
\end{align}
By calculating the thermodynamic forces for each of the components, which
are denoted by $\vect F^\typea_\thf$ and $\vect F^\typeb_\thf$,
we impose the correct density profile to the system:
\begin{align}
  \rho_\HY^\typea &= \rho_\AT^\typea = \rho_\CG^\typea = \rho_0^\typea\\
  \rho_\HY^\typeb &= \rho_\AT^\typeb = \rho_\CG^\typeb = \rho_0^\typeb
\end{align}
These can be achieved by the iterative thermodynamic force scheme for each
component in the mixture system.\\

Similar to Eq.~\eqref{eqn:peq}, we have for the force interpolation AdResS:
\begin{align}\label{eqn:peq-d-ab}
  p_\CG - p_\AT=\rho_0^\typea \omega^\typea_\thf + \rho_0^\typeb \omega^\typeb_\thf
\end{align}
Following the same argument in Sec.~\ref{sec:tmp4}, for $\typea$--$\typeb$
mixture we have
\begin{align}\label{eqn:peq-h-ab}
  p_\CG - p_\AT=
  \rho_0^\typea (\omega^{\typea,\hadress}_\thf - \omega^\typea_\res)
  +
  \rho_0^\typeb (\omega^{\typeb,\hadress}_\thf - \omega^\typeb_\res),
\end{align}
where the work of change representation for molecule $\typea$ is defined by
\begin{align}\label{eqn:wrep-split-a}
  \omega^\typea_\res 
  &=
  \omega^{\typea\typea}_\res +\omega^{\typea\typeb}_\res
  = \int_\HY d\vect r \langle \vect F^{\typea\typea}_\res(\vect r)\rangle
  + \int_\HY d\vect r \langle \vect F^{\typea\typeb}_\res(\vect r)\rangle,\\
  \label{eqn:wrep-split-b}
  \omega^\typeb_\res 
  &=
  \omega^{\typeb\typea}_\res +\omega^{\typeb\typeb}_\res
  = \int_\HY d\vect r \langle \vect F^{\typeb\typea}_\res(\vect r)\rangle
  + \int_\HY d\vect r \langle \vect F^{\typeb\typeb}_\res(\vect r)\rangle.
\end{align}
Here work of changing representation is splitted based on the
type of interaction, e.g.~$\omega_\res^{\typea\typeb}$ is due to the
cross interaction between molecule $\typea$ and $\typeb$.
We have the explicit expressions:
\begin{align}
  \langle \vect F^{\typea\typea}_\res(\vect r)\rangle
  &=
  \nabla_{\vect r} w(\vect r)\,
  \big\langle w(\vect r')
  \big[\, \sum_{\alpha,\beta} U^\AT_{\typea\typea}(\vect r_\alpha- \vect r'_\beta)
  - U^\CG_{\typea\typea}(\vect r- \vect r') \,\big]
  \big\rangle_{\vect r',\typea; \vect r,\typea}\\
  \langle \vect F^{\typea\typeb}_\res(\vect r)\rangle
  &=
  \nabla_{\vect r} w(\vect r)\,
  \big\langle w(\vect r')
  \big[\, \sum_{\alpha,\beta} U^\AT_{\typea\typeb}(\vect r_\alpha- \vect r'_\beta)
  - U^\CG_{\typea\typeb}(\vect r- \vect r') \,\big]
  \big\rangle_{\vect r',\typeb; \vect r,\typea}\\
  \langle \vect F^{\typeb\typea}_\res(\vect r)\rangle
  &=
  \nabla_{\vect r} w(\vect r)\,
  \big\langle w(\vect r')
  \big[\, \sum_{\alpha,\beta} U^\AT_{\typeb\typea}(\vect r_\alpha- \vect r'_\beta)
  - U^\CG_{\typeb\typea}(\vect r- \vect r') \,\big]
  \big\rangle_{\vect r',\typea; \vect r,\typeb}\\
  \langle \vect F^{\typeb\typeb}_\res(\vect r)\rangle
  &=
  \nabla_{\vect r} w(\vect r)\,
  \big\langle w(\vect r')
  \big[\, \sum_{\alpha,\beta} U^\AT_{\typeb\typeb}(\vect r_\alpha- \vect r'_\beta)
  - U^\CG_{\typeb\typeb}(\vect r- \vect r') \,\big]
  \big\rangle_{\vect r',\typeb; \vect r,\typeb}
\end{align}
The notations are self-explanatory. For example $ U^\AT_{\typea\typeb}$
denotes the atomistic interaction between molecule $\typea$ and $\typeb$.
Notation $\langle\cdot\rangle_{\vect r',\typea; \vect r,\typeb}$ denotes the
ensemble average taken with respect to position $\vect r'$ being molecule
$\typea$, provided a molecule $\typeb$ taking the postion $\vect r$.
If the molecules contain more than one atom, then the average is also taken
over all possible conformations.
Therefore, the physical meaning of the
force $\langle \vect F^{\typea\typeb}_\res(\vect r)\rangle$
is that providing a molecule of type $\typea$ at $\vect r$,
the averaged force between this atom and all molecule type $\typeb$
in the system.
Although we have $U^\AT_{\typea\typeb} = U^\AT_{\typeb\typea}$ and
$U^\CG_{\typea\typeb} = U^\CG_{\typeb\typea}$, it should
be noted that we do not have
$\omega_\res^{\typeb\typea} = \omega_\res^{\typea\typeb}$  in general.
From Eq.~\eqref{eqn:peq-d-ab}, \eqref{eqn:peq-h-ab},
\eqref{eqn:wrep-split-a} and \eqref{eqn:wrep-split-b}, we have
\begin{align}\label{eqn:hd-rel-ab}
  \rho_0^\typea(\omega_\thf^{\typea,\hadress} - \omega_\thf^\typea
  - \omega^{\typea\typea}_\res - \omega^{\typea\typeb}_\res)
  +
  \rho_0^\typeb(\omega_\thf^{\typeb,\hadress} - \omega_\thf^\typeb
  - \omega^{\typeb\typea}_\res - \omega^{\typeb\typeb}_\res) = 0
\end{align}

We denotes the work done by the transition region on the two types
of molecules by $\omega_0^\typea$ and $\omega_0^\typeb$, respectively.
To find out the chemical potential difference between AT and CG resolutions,
the same idea as presented in Sec.III.C of Ref.5 can be extended
to the two component system. This extension is straightforward and trivial. We do not want
to write them down, because one can just step-by-step follow the
deriviation for the one component system, and no new technique is needed.
In the end, we proved
\begin{align}\label{eqn:tmpc7}
  \mu^\typea_\AT(N^\typea_1, N^\typeb_1, V_1, T) &= \mu^\typea_\CG(N^\typea_1, N^\typeb_1, V_1, T) - \omega^\typea_0\\\label{eqn:tmpc8}
  \mu^\typeb_\AT(N^\typea_1, N^\typeb_1, V_1, T) &= \mu^\typeb_\CG(N^\typea_1, N^\typeb_1, V_1, T) - \omega^\typeb_0
\end{align}
Now, $N^\typea_1$ and $N^\typeb_1$ has the meaning of the averaged number of type $\typea$ and
type $\typeb$ molecules in the atomistic region, respectively.  Under the thermodynamic
limit, this number maximize the Helmholtz free energy and is the one
of statistical importance, see more discussions in
Ref.5. We do not distinguish these notations from
before, because the exact meaning is clear from the context.
The meaning of the chemical potential, e.g.~$\mu_\AT^\typea$, is clear:
it is the free energy increament of inserting one type $\typea$ molecule
into the infinitely large $\typea$--$\typeb$ mixture.\\
